### 2024-06-26 manuscript.Rmd
added missing information fixed some refs

factorFit1 <- mxRun(factorModel1)
summary(factorFit1)
```
Statistical models may often store internal information, which is not necessarily relevant to the statistical result but will strictly lead to a non-reproducibility error as fingerprints of the entire objects will differ.
For example, structural equation models estimated with `OpenMx` [@openmx2] store information about the time elapsed when fitting the model. 
Here is an example of the contents of a simple CFA model from the OpenMx documentation called `factorFit1` run on OpenMx's demonstration dataset `demoOneFactor` including 500 observations on five numeric variables. 
Among information relevant to reproducibility issues, its `output` attribute also contains timing variables `wallTime` and `cpuTime`, which will almost always lead to a non-reproducibility error if the entire `factorFit1` object is tested for reproducibility:

```{r eval=TRUE, echo=TRUE}
cat(factorFit1$output$wallTime)
cat(factorFit1$output$cpuTime)
```

Thus, we generally recommend a checkpoint approach, in which reproducibility tests are used only at selected, meaningful stages of the data analysis process and are used only on variables that contain values that are shown and/or interpreted in the scientific report (i.e., effect size point estimates, goodness-of-fit indices, standard error estimates, confidence intervals, test statistics, p values, Bayes factors, etc.).
Specifically, we recommend to use at least the following checkpoints:
- data loading: check whether the loaded raw data are identical to the data loaded in the original analysis
- preprocessing: check whether the preprocessed data (that is, after steps such as outlier removal, aggregation, filtering) are identical to the preprocessed data in the original analysis
- results: check whether the results that are reported in text, tables, and figures are identical to the results of the original analysis.
At the same time, avoid adding automated tests of entire statistical models but focus on the results.
A convenient approach is to use standard wrapper methods that extract relevant numeric quantities from statistical models, such as the generic function `coef()` which can be used to extract point estimates from linear regression models or `parTable()` in `lavaan` or `omxGetParameters()`in the aforementioned `OpenMx` model.
The output of the `summary()` function may be a good target for reproducibility checks because it often contains information about parameter estimates and fit statistics; however, in some cases (as with OpenMx models), it may again contain timing information, which will not exactly reproduce.
Last, note that some algorithm are not expected to reproduce identically but up to a certain precision.
This is relevant for algorithm that are fundamentally based on random numbers, such as Monte Carlo methods, bootstrap estimators and similar.
In this cases, approximate reproducibility tests can be realized by individually adjusting the numeric precision of the tests (see option `reproducibleRchunks.digits` as previously described).
Adepts of the tidyverse approach [@wickham2019welcome] are probably well advised to use functions `tidy()` and `glance()` from the `broom` [@broom] package, which will write parameter estimates and model fit statistics from various statistical models in R (e.g., anova, glm, coxph, gam, lm, lavaan,smooth.spline, survfit, and others) into a consistent and easy accessible format (a so-called _tibble_), which lends itself for fingerprinting with our suggested approach.
Here is a brief example of how the `broom` package formats the output of a linear regression model:

@@

@@

- Some computations are executed and their results are stored in a JSON file as planned. Later on, someone changes the code in the R Markdown file, such that the results differ and a failure of reproduction is indicated.

>>>>

factorFit1 <- mxRun(factorModel1)
summary(factorFit1)
```
Statistical models may often store internal information, which is not necessarily relevant to the statistical result but will strictly lead to a non-reproducibility error as fingerprints of the entire objects will differ.
For example, structural equation models estimated with `OpenMx` [@openmx2] store information about the time elapsed when fitting the model. 
Here is an example of the contents of a simple CFA model from the OpenMx documentation called `factorFit1` run on OpenMx's demonstration dataset `demoOneFactor` including 500 observations on five numeric variables. 
Among information relevant to reproducibility issues, its `output` attribute also contains timing variables `wallTime` and `cpuTime` as well as a timestamp `timestamp`, which will almost always lead to a non-reproducibility error if the entire `factorFit1` object is tested for reproducibility.
For illustration, here is the content of the `cpuTime` variable and the `timestump` variable for the aforementioned factor model, which was estimated when this manuscript was generated:

```{r eval=TRUE, echo=TRUE}
cat(factorFit1$output$cpuTime)
cat(factorFit1$output$timestamp)
```

Thus, we generally recommend a checkpoint approach, in which reproducibility tests are used only at selected, meaningful stages of the data analysis process and are used only on variables that contain values that are shown and/or interpreted in the scientific report (i.e., effect size point estimates, goodness-of-fit indices, standard error estimates, confidence intervals, test statistics, p values, Bayes factors, etc.).
Specifically, we recommend to use at least the following checkpoints:
- data loading: check whether the loaded raw data are identical to the data loaded in the original analysis
- preprocessing: check whether the preprocessed data (that is, after steps such as outlier removal, aggregation, filtering) are identical to the preprocessed data in the original analysis
- results: check whether the results that are reported in text, tables, and figures are identical to the results of the original analysis.
At the same time, avoid adding automated tests of entire statistical models but focus on the results.
A convenient approach is to use standard wrapper methods that extract relevant numeric quantities from statistical models, such as the generic function `coef()` which can be used to extract point estimates from linear regression models or `parTable()` in `lavaan` or `omxGetParameters()`in the aforementioned `OpenMx` model.
The output of the `summary()` function may be a good target for reproducibility checks because it often contains information about parameter estimates and fit statistics; however, in some cases (as with OpenMx models), it may again contain timing information, which will not exactly reproduce.
Last, note that some algorithm are not expected to reproduce identically but up to a certain precision.
This is relevant for algorithm that are fundamentally based on random numbers, such as Monte Carlo methods, bootstrap estimators and similar.
In this cases, approximate reproducibility tests can be realized by individually adjusting the numeric precision of the tests (see option `reproducibleRchunks.digits` as previously described).
Friends of the tidyverse approach [@wickham2019welcome] are probably well advised to use functions `tidy()` and `glance()` from the `broom` [@broom] package, which will write parameter estimates and model fit statistics from various statistical models in R (e.g., anova, glm, coxph, gam, lm, lavaan,smooth.spline, survfit, and others) into a consistent and easy accessible format (a so-called _tibble_), which lends itself for fingerprinting with our suggested approach.
Here is a brief example of how the `broom` package formats the output of a linear regression model:

@@

As a rule of thumb, we recommend to fingerprint the results of all parameter estimates from the `tidy()` function and all model fit statistics from the `glance()` function in a `reproducibleR` chunk.

@@

- Some computations are executed and their original results are stored in a JSON file as planned. Later on, someone modifies the R code in the R Markdown file, such that the results differ and a failure of reproduction is indicated. This change of code between the original computation and the reproduction attempt is caught by the package because fingerprints of the entire code chunk syntax are stored. Users are informed by a warning in the reproducibility report. The package will still try to reproduce each result and give individual reports on successes and failures.
